---
title: "Porcine data workflow"
output: html_document
params:
  repo_path:
    label: "Select the folder containing the repo on your local device"
    value: "C:/Users/anna-/OneDrive/Documents/Functional_Modelling_of_Melatonin_Action_in_the_Placenta"
    input: file
    type: directory
  
---

# Setup

```{r setup, include=FALSE}
#read in required packages
library(readr)
library(limma)
library(edgeR)
library(stringr)
library(pheatmap)
library(gplots)
library(tidyverse)
library(tidyr)
library(biomaRt)
library(dplyr)
library(umap)
library(Glimma)
library(RColorBrewer)
library(mclust)
library(rgl)
library(entropy)
library(pheatmap)

#file path to data
file_path <- file.path(params$repo_path, "porcine_data/porcine_counts.txt")

#Read in data
data <- read.table(file_path, header=TRUE)

#set row names to GeneID
rownames(data) <- data$Geneid # sets row names to GeneID

```

# Mapping the Gene names using the Ensembl database

``` {r Map gene names using the ensembl database}

Geneid <- data$Geneid #Extract all gene names

#Connect to porcine Ensembl dataset (use archived host to avoid server error):
ensembldatabase <- useMart(biomart="ensembl", dataset = "sscrofa_gene_ensembl", verbose=TRUE, host="https://feb2021.archive.ensembl.org")

#Set up a query to convert the gene names (use getBM when only using one database, get LDS for two):
gene_names <- getBM(attributes = c("ensembl_gene_id", "external_gene_name"),
                    filters ="ensembl_gene_id",
                    values = Geneid,
                    mart = ensembldatabase)

```

``` {r formatting mapped gene names}

#GENENAME = external name
#GENEID = ensembl ID
colnames(gene_names)[1] <- 'Geneid'
colnames(gene_names)[2] <- 'GENENAME'

gene_names$GENENAME <-as.character(gene_names$GENENAME)
gene_names$GENENAME <- make.unique(gene_names$GENENAME,sep = "~")

#Join data to new gene_names
data1 <- data %>% left_join(gene_names, by = c("Geneid"))

#check there are no NAs
any(is.na(data1$Geneid))

genes <- data1$GENENAME# takes first column as gene name
rownames(data1) <- make.unique(genes,sep = "~") #reimput geneID as row names. This will add .1 to the end of gene names


